R Notebook

### 2020-02-12 src/testing_file.Rmd
updated plots (background point gen)

#get max and min depth layers based on max kappa
```{r}
kapd <- list.files("../output/maxent/paper_plots/prob_by_depth/" , pattern = "_maxkap_dis_prob_cells.csv", full.names = TRUE)

test <- read.csv(kapd[1])
test <- subset(test, Probability == 1) #just keep the presence cells
test <- test[order(test[, 1]),] #order by depth 
#now need to find the row in which a value first appears in prob_bin column

@@

>>>>

#get max and min depth layers based on max kappa
```{r}
kapd <- list.files("../output/maxent/paper_plots/prob_by_depth/" , pattern = "_maxkap_dis_prob_cells.csv", full.names = TRUE)
kapmmd <- data.frame(mth = character(10), min = numeric(10), max = numeric(10))
kapmmd$mth <- as.character(kapmmd$mth)

for(k in 1:length(kapd)){
  kapmmd$mth[k] <- str_sub(kapd[k], start = 44, end = 46)
  kd <- read.csv(kapd[k])
  kd <- subset(kd, Probability == 1) #just keep the presence cells
  kd <- subset(kd, prob_bin > 0) # get the depths where there are presence cells
  kd <- kd[order(kd[, 1]),] #order by depth 
  kapmmd$min[k] <- kd$Depth[1] #minimum depth of presence cell
  kapmmd$max[k] <- tail(kd$Depth, n = 1) #maximum depth of presence cell
}

write.csv(kapmmd, "../output/maxent/paper_plots/hotspot_mapping/kappa-min-max-depth.csv")
```

#get lat long position of most northely, southerly, western, and eastern cell centroid

```{r}
filelist <- list.files("../output/maxent/paper_plots/hotspot_mapping/", pattern = "maxkap_single.asc", full.names = TRUE)
lonlat <- data.frame(mth = character(10), west = numeric(10), east = numeric(10), north = numeric(10), south = numeric(10))
lonlat$mth <- as.character(lonlat$mth)

for (k in 1:length(filelist)){
  ras <- raster(filelist[9])
  crs(ras) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" #the projection of the raster
  lonlat$mth[k] <- str_sub(filelist[k], start = 46, end = 48)
  cellloc <- rasterToPoints(ras, fun=function(x){x==1}) #make raster to point only for cells with the value of 1 (i.e. presence cells)
  cellloc <- as.data.frame(cellloc) # x = lon, y = lat
  cellloc <- cellloc[order(cellloc$x), ] #order dataframe by lon from low (west) to high (east)
  lonlat$west[k] <- cellloc$x[1] #most western cell centroid
  lonlat$east[k] <- tail(cellloc$x, n = 1) #most eastern cell centroid
  cellloc <- cellloc[order(cellloc$y), ] #order dataframe by lat from low (south) to high (north)
  lonlat$south[k] <- cellloc$y[1] #most southern cell centroid
  lonlat$north[k] <- tail(cellloc$y, n = 1) #most northern cell centroid
}

write.csv(lonlat, "../output/maxent/paper_plots/hotspot_mapping/kappa-nswe.csv")

@@

#get lat long position of most northely, southerly, western, and eastern cell centroid for oct, june, aug, may (to remove northern spots)

```{r}
filelist <- list.files("../output/maxent/paper_plots/hotspot_mapping/", pattern = "maxkap_single.asc", full.names = TRUE)
lonlat <- data.frame(mth = character(10), west = numeric(10), east = numeric(10), north = numeric(10), south = numeric(10))
lonlat$mth <- as.character(lonlat$mth)

for (k in 1:length(filelist)){
  ras <- raster(filelist[k])
  crs(ras) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" #the projection of the raster
  lonlat$mth[k] <- str_sub(filelist[k], start = 46, end = 48)
  cellloc <- rasterToPoints(ras, fun=function(x){x==1}) #make raster to point only for cells with the value of 1 (i.e. presence cells)
  cellloc <- as.data.frame(cellloc) # x = lon, y = lat
  cellloc <- subset(cellloc, y <55) #remove all the northern points
  cellloc <- cellloc[order(cellloc$y), ] #order dataframe by lat from low (south) to high (north)
  lonlat$south[k] <- cellloc$y[1] #most southern cell centroid
  lonlat$north[k] <- tail(cellloc$y, n = 1) #most northern cell centroid
  cellloc <- cellloc[order(cellloc$x), ] #order dataframe by lon from low (west) to high (east)
  lonlat$west[k] <- cellloc$x[1] #most western cell centroid
  lonlat$east[k] <- tail(cellloc$x, n = 1) #most eastern cell centroid

write.csv(lonlat, "../output/maxent/paper_plots/hotspot_mapping/kappa-nswe_nonorth.csv")

```{r}
t1 <- distVincentyEllipsoid(c(0,0),c(90,90))
# on a 'Clarke 1880' ellipsoid
test <- distVincentyEllipsoid(c(0,0),c(90,90), a=6378249.145, b=6356514.86955, f=1/293.465)
```

#background point selection plots

```{r}
test_back_df <- read.csv("../data/env/background_point_check/201406_allbackgroundpoints.csv", header = TRUE)
```

ok now create the first lot of random for 2014_06

```{r}
testbk10000 <- test_back_df[sample(nrow(test_back_df), 10000), ]  #where 10000 = number of rows to sample (large sample as per maxent)
testbk20000 <- test_back_df[sample(nrow(test_back_df), 20000), ]  
testbk50000 <- test_back_df[sample(nrow(test_back_df), 50000), ]  
testbk100000 <- test_back_df[sample(nrow(test_back_df), 100000), ]  
testbk190000 <- test_back_df[sample(nrow(test_back_df), 190000), ]  
```

make long data
```{r}
bkpt <- data.frame(no_points = character(370000), variable = character(370000), value = numeric(370000))
bkpt$variable <- as.character(bkpt$variable)
bkpt$no_points <- as.character(bkpt$no_points)

bkpt$no_points[1:10000] <- "10,000"
bkpt$no_points[10001:30000] <- "20,000"
bkpt$no_points[30001:80000] <- "50,000"
bkpt$no_points[80001:180000] <- "100,000"
bkpt$no_points[180001:370000] <- "190,000"
bkpt$variable <- "temp"
bkpt$value[1:10000] <- testbk10000$temp_depth
bkpt$value[10001:30000] <- testbk20000$temp_depth
bkpt$value[30001:80000] <- testbk50000$temp_depth
bkpt$value[80001:180000] <- testbk100000$temp_depth
bkpt$value[180001:370000] <- testbk190000$temp_depth

bkpt2 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
bkpt2$variable <- as.character(bkpt2$variable)
bkpt2$no_points[1:10000] <- "10,000"
bkpt2$no_points[10001:30000] <- "20,000"
bkpt2$no_points[30001:80000] <- "50,000"
bkpt2$no_points[80001:180000] <- "100,000"
bkpt2$no_points[180001:370000] <- "190,000"
bkpt2$variable <- "salinity"
bkpt2$value[1:10000] <- testbk10000$salinity_depth
bkpt2$value[10001:30000] <- testbk20000$salinity_depth
bkpt2$value[30001:80000] <- testbk50000$salinity_depth
bkpt2$value[80001:180000] <- testbk100000$salinity_depth
bkpt2$value[180001:370000] <- testbk190000$salinity_depth

bkpt3 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
bkpt3$variable <- as.character(bkpt3$variable)
bkpt3$no_points[1:10000] <- "10,000"
bkpt3$no_points[10001:30000] <- "20,000"
bkpt3$no_points[30001:80000] <- "50,000"
bkpt3$no_points[80001:180000] <- "100,000"
bkpt3$no_points[180001:370000] <- "190,000"
bkpt3$variable <- "o2"
bkpt3$value[1:10000] <- testbk10000$o2_depth
bkpt3$value[10001:30000] <- testbk20000$o2_depth
bkpt3$value[30001:80000] <- testbk50000$o2_depth
bkpt3$value[80001:180000] <- testbk100000$o2_depth
bkpt3$value[180001:370000] <- testbk190000$o2_depth

bkpt4 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
bkpt4$variable <- as.character(bkpt4$variable)
bkpt4$no_points[1:10000] <- "10,000"
bkpt4$no_points[10001:30000] <- "20,000"
bkpt4$no_points[30001:80000] <- "50,000"
bkpt4$no_points[80001:180000] <- "100,000"
bkpt4$no_points[180001:370000] <- "190,000"
bkpt4$variable <- "chl"
bkpt4$value[1:10000] <- testbk10000$chl_surface
bkpt4$value[10001:30000] <- testbk20000$chl_surface
bkpt4$value[30001:80000] <- testbk50000$chl_surface
bkpt4$value[80001:180000] <- testbk100000$chl_surface
bkpt4$value[180001:370000] <- testbk190000$chl_surface

bk <- rbind(bkpt, bkpt2, bkpt3, bkpt4)
bk$no_points <- as.factor(bk$no_points)
```

```{r}
ggplot(bk, aes(x = value)) +
    stat_density(aes(group = no_points, color = no_points),position="identity",geom="line") +
  facet_wrap(~variable,ncol=2)

problem here is that i don't want the axis shared

fiddle with the dataset again
```{r}
t <- subset(bk, variable == "temp")
t$no_points <- factor(t$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
s <- subset(bk, variable == "salinity")
s$no_points <- factor(s$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
o <- subset(bk, variable == "o2")
o$no_points <- factor(o$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
c <- subset(bk, variable == "chl")
c$no_points <- factor(c$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))

plot
```{r}
p1 <- ggplot(t, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) +theme_classic(base_size = 18) + labs(x = "Temperature (Kelvin)", y = "Density") + guides(color = guide_legend(title = "No Points"))
p2 <- ggplot(s, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) + theme_classic(base_size = 18) + labs (x = "Salinity (PSU)", y = "Density") + guides(color = guide_legend(title = "No Points"))
p3 <- ggplot(o, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) +theme_classic(base_size = 18) + labs (x = "Dissolved Oxygen (mmol\u207B\u00B3)", y = "Density") + guides(color = guide_legend(title = "No Points"))
p4 <- ggplot(c, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) + theme_classic(base_size = 18) + labs (x = "Chlorophyll \u03b1  (mmol\u207B\u00B3)", y = "Density") + guides(color = guide_legend(title = "No Points"))

p <- ggarrange(p1, p2, p3, p4, ncol = 2, nrow = 2, common.legend = TRUE, legend = "bottom")

ggsave("../output/maxent/paper_plots/bkpt_201406.png", width = 210, height = 297, units = "mm", dpi = 600)

200710

```{r}
test_back_df <- read.csv("../data/env/background_point_check/200710_allbackgroundpoints.csv", header = TRUE)
```

ok now create the first lot of random for 200710

```{r}
testbk10000 <- test_back_df[sample(nrow(test_back_df), 10000), ]  #where 10000 = number of rows to sample (large sample as per maxent)
testbk20000 <- test_back_df[sample(nrow(test_back_df), 20000), ]  
testbk50000 <- test_back_df[sample(nrow(test_back_df), 50000), ]  
testbk100000 <- test_back_df[sample(nrow(test_back_df), 100000), ]  
testbk190000 <- test_back_df[sample(nrow(test_back_df), 190000), ]  
```

make long data
```{r}
bkpt <- data.frame(no_points = character(370000), variable = character(370000), value = numeric(370000))
bkpt$variable <- as.character(bkpt$variable)
bkpt$no_points <- as.character(bkpt$no_points)

bkpt$no_points[1:10000] <- "10,000"
bkpt$no_points[10001:30000] <- "20,000"
bkpt$no_points[30001:80000] <- "50,000"
bkpt$no_points[80001:180000] <- "100,000"
bkpt$no_points[180001:370000] <- "190,000"
bkpt$variable <- "temp"
bkpt$value[1:10000] <- testbk10000$temp_depth
bkpt$value[10001:30000] <- testbk20000$temp_depth
bkpt$value[30001:80000] <- testbk50000$temp_depth
bkpt$value[80001:180000] <- testbk100000$temp_depth
bkpt$value[180001:370000] <- testbk190000$temp_depth

bkpt2 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
bkpt2$variable <- as.character(bkpt2$variable)
bkpt2$no_points[1:10000] <- "10,000"
bkpt2$no_points[10001:30000] <- "20,000"
bkpt2$no_points[30001:80000] <- "50,000"
bkpt2$no_points[80001:180000] <- "100,000"
bkpt2$no_points[180001:370000] <- "190,000"
bkpt2$variable <- "salinity"
bkpt2$value[1:10000] <- testbk10000$salinity_depth
bkpt2$value[10001:30000] <- testbk20000$salinity_depth
bkpt2$value[30001:80000] <- testbk50000$salinity_depth
bkpt2$value[80001:180000] <- testbk100000$salinity_depth
bkpt2$value[180001:370000] <- testbk190000$salinity_depth

bkpt3 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
bkpt3$variable <- as.character(bkpt3$variable)
bkpt3$no_points[1:10000] <- "10,000"
bkpt3$no_points[10001:30000] <- "20,000"
bkpt3$no_points[30001:80000] <- "50,000"
bkpt3$no_points[80001:180000] <- "100,000"
bkpt3$no_points[180001:370000] <- "190,000"
bkpt3$variable <- "o2"
bkpt3$value[1:10000] <- testbk10000$o2_depth
bkpt3$value[10001:30000] <- testbk20000$o2_depth
bkpt3$value[30001:80000] <- testbk50000$o2_depth
bkpt3$value[80001:180000] <- testbk100000$o2_depth
bkpt3$value[180001:370000] <- testbk190000$o2_depth

bkpt4 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
bkpt4$variable <- as.character(bkpt4$variable)
bkpt4$no_points[1:10000] <- "10,000"
bkpt4$no_points[10001:30000] <- "20,000"
bkpt4$no_points[30001:80000] <- "50,000"
bkpt4$no_points[80001:180000] <- "100,000"
bkpt4$no_points[180001:370000] <- "190,000"
bkpt4$variable <- "chl"
bkpt4$value[1:10000] <- testbk10000$chl_surface
bkpt4$value[10001:30000] <- testbk20000$chl_surface
bkpt4$value[30001:80000] <- testbk50000$chl_surface
bkpt4$value[80001:180000] <- testbk100000$chl_surface
bkpt4$value[180001:370000] <- testbk190000$chl_surface

bk <- rbind(bkpt, bkpt2, bkpt3, bkpt4)
bk$no_points <- as.factor(bk$no_points)
```

fiddle with the dataset again
```{r}
t <- subset(bk, variable == "temp")
t$no_points <- factor(t$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
s <- subset(bk, variable == "salinity")
s$no_points <- factor(s$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
o <- subset(bk, variable == "o2")
o$no_points <- factor(o$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
c <- subset(bk, variable == "chl")
c$no_points <- factor(c$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))

plot
```{r}
p1 <- ggplot(t, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) +theme_classic(base_size = 18) + labs(x = "Temperature (Kelvin)", y = "Density") + guides(color = guide_legend(title = "No Points"))
p2 <- ggplot(s, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) + theme_classic(base_size = 18) + labs (x = "Salinity (PSU)", y = "Density") + guides(color = guide_legend(title = "No Points"))
p3 <- ggplot(o, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) +theme_classic(base_size = 18) + labs (x = "Dissolved Oxygen (mmol\u207B\u00B3)", y = "Density") + guides(color = guide_legend(title = "No Points"))
p4 <- ggplot(c, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) + theme_classic(base_size = 18) + labs (x = "Chlorophyll \u03b1  (mmol\u207B\u00B3)", y = "Density") + guides(color = guide_legend(title = "No Points"))

p <- ggarrange(p1, p2, p3, p4, ncol = 2, nrow = 2, common.legend = TRUE, legend = "bottom")

ggsave("../output/maxent/paper_plots/bkpt_200710.png", width = 210, height = 297, units = "mm", dpi = 600)

```
199902

```{r}
test_back_df <- read.csv("../data/env/background_point_check/199902_allbackgroundpoints.csv", header = TRUE)
```

ok now create the first lot of random for 200710

```{r}
testbk10000 <- test_back_df[sample(nrow(test_back_df), 10000), ]  #where 10000 = number of rows to sample (large sample as per maxent)
testbk20000 <- test_back_df[sample(nrow(test_back_df), 20000), ]  
testbk50000 <- test_back_df[sample(nrow(test_back_df), 50000), ]  
testbk100000 <- test_back_df[sample(nrow(test_back_df), 100000), ]  
testbk190000 <- test_back_df[sample(nrow(test_back_df), 190000), ]  
```

make long data
```{r}
bkpt <- data.frame(no_points = character(370000), variable = character(370000), value = numeric(370000))
bkpt$variable <- as.character(bkpt$variable)
bkpt$no_points <- as.character(bkpt$no_points)

bkpt$no_points[1:10000] <- "10,000"
bkpt$no_points[10001:30000] <- "20,000"
bkpt$no_points[30001:80000] <- "50,000"
bkpt$no_points[80001:180000] <- "100,000"
bkpt$no_points[180001:370000] <- "190,000"
bkpt$variable <- "temp"
bkpt$value[1:10000] <- testbk10000$temp_depth
bkpt$value[10001:30000] <- testbk20000$temp_depth
bkpt$value[30001:80000] <- testbk50000$temp_depth
bkpt$value[80001:180000] <- testbk100000$temp_depth
bkpt$value[180001:370000] <- testbk190000$temp_depth

bkpt2 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
bkpt2$variable <- as.character(bkpt2$variable)
bkpt2$no_points[1:10000] <- "10,000"
bkpt2$no_points[10001:30000] <- "20,000"
bkpt2$no_points[30001:80000] <- "50,000"
bkpt2$no_points[80001:180000] <- "100,000"
bkpt2$no_points[180001:370000] <- "190,000"
bkpt2$variable <- "salinity"
bkpt2$value[1:10000] <- testbk10000$salinity_depth
bkpt2$value[10001:30000] <- testbk20000$salinity_depth
bkpt2$value[30001:80000] <- testbk50000$salinity_depth
bkpt2$value[80001:180000] <- testbk100000$salinity_depth
bkpt2$value[180001:370000] <- testbk190000$salinity_depth

bkpt3 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
bkpt3$variable <- as.character(bkpt3$variable)
bkpt3$no_points[1:10000] <- "10,000"
bkpt3$no_points[10001:30000] <- "20,000"
bkpt3$no_points[30001:80000] <- "50,000"
bkpt3$no_points[80001:180000] <- "100,000"
bkpt3$no_points[180001:370000] <- "190,000"
bkpt3$variable <- "o2"
bkpt3$value[1:10000] <- testbk10000$o2_depth
bkpt3$value[10001:30000] <- testbk20000$o2_depth
bkpt3$value[30001:80000] <- testbk50000$o2_depth
bkpt3$value[80001:180000] <- testbk100000$o2_depth
bkpt3$value[180001:370000] <- testbk190000$o2_depth

bkpt4 <- data.frame(no_points = numeric(370000), variable = character(370000), value = numeric(370000))
bkpt4$variable <- as.character(bkpt4$variable)
bkpt4$no_points[1:10000] <- "10,000"
bkpt4$no_points[10001:30000] <- "20,000"
bkpt4$no_points[30001:80000] <- "50,000"
bkpt4$no_points[80001:180000] <- "100,000"
bkpt4$no_points[180001:370000] <- "190,000"
bkpt4$variable <- "chl"
bkpt4$value[1:10000] <- testbk10000$chl_surface
bkpt4$value[10001:30000] <- testbk20000$chl_surface
bkpt4$value[30001:80000] <- testbk50000$chl_surface
bkpt4$value[80001:180000] <- testbk100000$chl_surface
bkpt4$value[180001:370000] <- testbk190000$chl_surface

bk <- rbind(bkpt, bkpt2, bkpt3, bkpt4)
bk$no_points <- as.factor(bk$no_points)
```

fiddle with the dataset again
```{r}
t <- subset(bk, variable == "temp")
t$no_points <- factor(t$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
s <- subset(bk, variable == "salinity")
s$no_points <- factor(s$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
o <- subset(bk, variable == "o2")
o$no_points <- factor(o$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))
c <- subset(bk, variable == "chl")
c$no_points <- factor(c$no_points, levels = c("10,000", "20,000", "50,000", "100,000", "190,000"))

plot
```{r}
p1 <- ggplot(t, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) +theme_classic(base_size = 18) + labs(x = "Temperature (Kelvin)", y = "Density") + guides(color = guide_legend(title = "No Points"))
p2 <- ggplot(s, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) + theme_classic(base_size = 18) + labs (x = "Salinity (PSU)", y = "Density") + guides(color = guide_legend(title = "No Points"))
p3 <- ggplot(o, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) +theme_classic(base_size = 18) + labs (x = "Dissolved Oxygen (mmol\u207B\u00B3)", y = "Density") + guides(color = guide_legend(title = "No Points"))
p4 <- ggplot(c, aes(x = value)) + stat_density(aes(group = no_points, color = no_points),position="identity",geom="line", size=0.75) + theme_classic(base_size = 18) + labs (x = "Chlorophyll \u03b1  (mmol\u207B\u00B3)", y = "Density") + guides(color = guide_legend(title = "No Points"))

p <- ggarrange(p1, p2, p3, p4, ncol = 2, nrow = 2, common.legend = TRUE, legend = "bottom")

ggsave("../output/maxent/paper_plots/bkpt_199902.png", width = 210, height = 297, units = "mm", dpi = 600)
